---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
num_variables <- ncol(mtcars)
num_observations <- nrow(mtcars)

# Answer:
print(paste("There are total of", num_variables, "variables and", num_observations, "observations in this data set."))
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary_stats <- summary(mtcars)

discrete_variables <- c("cyl", "vs", "am", "gear", "carb")
variables <- colnames(mtcars)
continuous_variables <- setdiff(variables, discrete_variables)

num_discrete_variables <- length(discrete_variables)
num_continuous_variables <- length(continuous_variables)

# Answer:
print(paste("There are total of", discrete_variables, "variables and", num_continuous_variables, "observations in this data set."))
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
mpg_data <- mtcars$mpg
m <- mean(mpg_data)
v <- var(mpg_data)
s <- sd(mpg_data)


print(paste("The average of Mile Per Gallon from this data set is", m, "with variance", v, "and standard deviation", s, "."))

```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
mpg_by_cyl <- tapply(mtcars$mpg, mtcars$cyl, mean)
mpg_by_gear <- tapply(mtcars$mpg, mtcars$gear, sd)

mpg_by_cyl_df <- data.frame(Cylinder_Class = as.numeric(names(mpg_by_cyl)), Average_MPG = mpg_by_cyl)
mpg_by_gear_df <- data.frame(Gear_Class = as.numeric(names(mpg_by_gear)), Standard_Deviation_MPG = mpg_by_gear)

print("1) Average MPG for each cylinder class:")
print(mpg_by_cyl_df)

print("\n2) Standard Deviation of MPG for each gear class:")
print(mpg_by_gear_df)
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
crosstab <- table(mtcars$cyl, mtcars$gear)
max_combination <- which.max(crosstab)
max_count <- crosstab[max_combination]

print("Combination with the most observations:")
print(paste("Cylinder Class: ", rownames(crosstab)[max_combination[1]]))
print(paste("Gear Class: ", colnames(crosstab)[max_combination[2]]))
print(paste("Number of Observations: ", max_count))

print(paste("The combination with the most observations is", rownames(crosstab)[max_combination[1]], "cylinders and", colnames(crosstab)[max_combination[2]], "gears, with", max_count, "observations."))
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
boxplot(weight ~ group, data = PlantGrowth,
        main = "Weight of Plants by Group",
        xlab = "Group",
        ylab = "Weight")

group_labels <- levels(PlantGrowth$group)

```

Result:

=> Report a paragraph to summarize your findings from the plot!



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg, breaks = 10, main = "MPG Histogram", xlab = "Miles Per Gallon (MPG)", ylab = "Frequency")
hist_data <- hist(mtcars$mpg, breaks = 10, plot = FALSE)
most_observed_mpg_class <- hist_data$mids[which.max(hist_data$counts)]
abline(v = most_observed_mpg_class, col = "red")


print(paste("Most of the cars in this data set are in the class of", most_observed_mpg_class, "mile per gallon."))
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests, main = "Pairs Plot of USArrests Dataset")
plot(USArrests$Murder, USArrests$Assault, main = "Scatter Plot of Murder vs. Assault", xlab = "Murder", ylab = "Assault")
```

Result:

=> Report a paragraph to summarize your findings from the plot!



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
market_value_stats <- summary(housingData$Market.Value.per.SqFt)
agg_table <- aggregate(Market.Value.per.SqFt ~ Boro, data = housingData, FUN = mean)
year_built_counts <- table(housingData$Year.Built)

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
plot(housingData$Year.Built, housingData$Market.Value.per.SqFt, xlab = "Year Built", ylab = "Market Value per SqFt", main = "Scatter Plot: Market.Value.per.SqFt vs. Year.Built")

par(mar = c(5, 10, 4, 2))
plot(housingData$Neighborhood, housingData$Market.Value.per.SqFt, xlab = "Neighborhood", ylab = "Market Value per SqFt", main = "Scatter Plot: Market.Value.per.SqFt vs. Neighborhood")
axis(1, at = 1:nlevels(factor(housingData$Neighborhood)), labels = levels(factor(housingData$Neighborhood)), las = 2)
```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!


